\section{Conclusions} \label{sec:conclusions}

This work investigated the hazards in software masking, suggested a verification tool and established a secure, 1st-order masked Sbox implementation against 1st-order, univariate attacks. Still, several important questions for future work arise. We demonstrated that removing the ILA-breaching effects is feasible, yet identifying the best clearing mechanism and minimizing the overhead is a topic for further exploration. Similarly, the current work is limited in AVR ATMega163 and needs to be extended to different devices and platforms. It could be done by using ASCOLD tool as a base for this kind of work.  Moreover, higher-order evaluation techniques are still nascent and in this work we did not focus on 1st-order, yet multivariate attacks such as those that exploit horizontality~\cite{DBLP:conf/ches/BattistelloCPZ16}. In addition, note that the ILA effects are observable throughout an implementation. Not only the cipher-related operations but any manipulation of shares during I/O, RNG routines etc. can create hazards. Thus, there is need for effort towards a fully hardened implementation. Last but not least, we stress that the effects identified depend on the architecture and the physical layer, thus preventing them in the assembly layer is, in principle, less efficient and prone to errors. Future work can strive towards custom-made microcontrollers that enforce ILA in hardware. Ideally, such a microcontroller should be able to guarantee ILA without additional countermeasures such as threshold implementations~\cite{DBLP:conf/icics/NikovaRR06}.  